Insights of Even-point Multi-loop Unitarity: Exponentiation, Anomalies and Evanescence

### 2023-06-14 eventPointTwoLoop.tex
citations
added citations fixed typos

When expressed in this form, many gauge theories reveal hidden redundancy that is obscured by the color-dressed amplitude formulation of \eqn{fullAmp}. When expressed in this form, the partial amplitudes for large catalog \cite{BCJreview} of gauge theories, including Yang-Mills and NLSM, are linearly dependent. The first such set of partial amplitude relations identified by Kleiss and Kluijf \cite{KK}, known formally as KK relations, relates amplitudes different color orderings and signature:
\be\label{KK}
A(1,\alpha,n,\beta)=(-1)^{|\alpha|}\sum_{\sigma\in \alpha \shuffle \beta^T} A(1,\sigma,n)
\ee
where $\beta^T$ denotes the inverse ordering of $\beta$, and $\shuffle$ is the shuffle product that enumerates ordered permutations of $\alpha$ and $\beta^T$. More recently, Bern, Johansson and one of the authors, further identified a set of kinematic relations that reduced the size of $n$-point partial amplitude basis down to $(n-3)!$ via BCJ relations \cite{BCJ}:
\be\label{BCJ}
\sum_{i=2}^{n-1}k_1\cdot(k_2+\cdots +k_i) A(2,...,i,1,...,n)=0
\ee
Underlying both sets of amplitude relations are hidden graphical organization principles that manifest the amplitude level redundancy of \eqn{KK} and \eqn{BCJ}, which we will describe in the next section. Before moving on, we emphasize that everything described thus far applies in arbitrary spacetime dimensions. As such, the aforementioned nomenclature and organization is well suited for constructing loop integrands needed for dimensional regularization.
\label{subsec:AmpReview}

@@

Formal polarizations will always be expressed as $\varepsilon_a^\mu$ to distinguish between the extra dimensional bit of the integration measure in $D=4-2\epsilon$. As noted, it will be convenient to take $D\rightarrow 4$ in certain circumstances after integration. We will use the conventions of ref. \cite{jjmcTASI2014} throughout the text, which we provide below

@@

We could also consider higher derivative generalizations of the NLSM that are invariant under the shift symmetry of \eqn{eq:shiftSym}. Such EFTs are known formally as \textit{chiral perturbation theory}, and descriptors of pion scattering below the chiral symmetry breaking scale of QCD. 

@@

By construction, each of the higher derivative operators appearing above are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{}, and soft theorem bootstraps of refs. \cite{}. Considering the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $SU(2)$ NLSM have been computed through two-loop order \cite{} . In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. In doing so, we show that the IR divergences of the theory exponentiates in the large $N_c$ limit. 

@@

The second even-point EFT we will study is Born-Infeld theory \cite{}. This theory describes the dynamics of open string endpoints with $U(1)$ charge attached to a $D$-dimensional spacetime with Dirichlet boundary conditions. Given the stringy dynamics orthogonal to the spacetime, the electric field generated by the charged endpoints has a maximum field strength of order the inverse string tension, $1/\alpha'$. The effective Lagrangian for this theory

@@

where $\varphi$ is the spacetime coordinate in the orthogonal space. This theory is uniquely determined by requiring the a polynomial EFT of $X=(\partial \varphi)^2$ is invariant under the field redefinition:

@@

When the transformation law leave the polynomial $P(X)$ theory invariant, there are an additional $D+1$ conserved charges, and the Poincar\'{e} symmetry of $P(X)$ EFT is promoted from $D$ to $(D+1)$ dimensions. One can verify that applying \eqn{DBIShift} to \eqn{DBI} shifts the Lagrangian by a total derivative. Similar to the soft theorem constraints on the building blocks of NLSM and $\chi \text{PT}$, the DBI action can be bootstrapped by requiring the amplitudes vanish at $\mathcal{O}(p^2)$ when taking external momentum, $p$, soft. 

@@

Now we will describe the supersymmetric extensions of the Born-Infeld action given above in \eqn{biLag}. First we start by defining the theory of shift symmetric fermions first written down by Volkov, Akulov (VA) \cite{}. The easiest way to define this shift symmetric theory is by first constructing manifestly symmetric 1-forms, and constructing a volume form in the veirbein frame. That is, define a 1-form, $ \omega^m = e^m_\mu dx^\mu$, where the veirbein (frame fields), $e^m_\mu$, is expressed in terms of the fermion fields, $\lambda$, as follows:
 \be
 e^m_\mu = \delta^m_\mu + i \bar{\lambda} \Gamma^m\! \stackrel{\leftrightarrow}{\partial_\mu}\!  \lambda = \delta^m_\mu + i (\bar{\lambda} \Gamma^m \partial_\mu  \lambda - \partial_\mu \bar{\lambda} \Gamma^m   \lambda)
 \ee
 This frame field is manifestly invariant under the superspace shift:
\be
\delta \lambda \rightarrow \eta \qquad  \delta \bar{\lambda} \rightarrow \bar{\eta} \qquad  \delta x^\mu \rightarrow x^\mu - i (\bar{\lambda} \,\Gamma^\mu \eta- i \bar{\eta} \,\Gamma^\mu \lambda)
\ee
 The Volkov-Akulov action in the veirbein frame is then simply the fermionic world-volume integral over the $D$-form:
 \be
 S^{\text{VA}} = \int \omega^1 \wedge \omega^2 \wedge \cdots \wedge \omega^D
 \ee
 This is not dissimilar from the construction of the pure scale DBI action above. Likewise, we can write the Lagrangian in the spacetime coordinate from akin to \eqn{biLag} and \eqn{DBI}. Transforming back into spacetime coordinates, yields the following Lagrangian density for self interaction VA fermions upto four-point interactions:
\begin{equation}
 \label{vaLag}
  \alpha'^{\,2}\mathcal{L}^{\text{VA}} = \text{det}(e^m_\mu) =  i\bar{\lambda}\!\stackrel{\leftrightarrow}{\slashed{\partial}}\!\lambda+ \frac{1}{2}(\bar{\lambda}\Gamma^\mu \partial^\nu \lambda)(\bar{\lambda}\Gamma_\mu \partial_\nu \lambda)+\mathcal{O}(\partial^4\lambda^6) 
  \,.
\end{equation}
 A convenient choice of field redefinition was proposed by Komogorov and Seiberg (KS) \cite{}, which defines a new fermion field, $\psi = \psi(\lambda,\bar{\lambda})$, in a way that is perturbatively equivalent to the VA fermions. This construction is equivalent to a theory of nilpotent chiral superfields \cite{}, relevant for the construction of $\alpha$-attractor models of inflation \cite{}. Applying the KS nonlinear field redefinition yields the considerably simpler Lagrangian for VA theory:
 \be
   \alpha'^{\,2}\mathcal{L}^{\text{VA}} = i\bar{\psi}\slashed{\partial}\psi +\frac{1}{2} \bar{\psi}^2 \partial^2 \psi^2 + \frac{1}{4}\psi^2\bar{\psi}^2 \partial^2 \psi^2\partial^2 \bar{\psi}^2
 \ee
With the world-volume formulation of VA theory in hand, we can also trivially construct the action for maximal $\mathcal{N}=1$ supersymmetric Born-Infeld theory in $D=10$. We simply replace the flat space background of \eqn{biLag} with the fermionic background of \eqn{vaLag},
\be
 \label{DBIVALag}
S^{\text{DBIVA}}_{\mathcal{N}=1} = \int \omega^1 \wedge\omega^2 \wedge \cdots \wedge \omega^{10} \sqrt{\text{det}(\eta_{mn}+ \alpha F_{mn})}
\ee
It is worth noting that the fermion-vector interaction is introduced via the \textit{non-minimal} coupling between the field strength and the fermionic spacetime metric. This differs from the linearly realized supersymmetry of super-Yang-Mills theory, which minimally couples the vector mode to the chiral fermion. This non-minimal coupling is a requirement of supersymmtry due to the even-point nature abelian vector theories like Born-Infeld.

@@

Furthermore, factors of $(\epsilon_i\cdot l)$ can be mapped to tensor structures with factors of $(\epsilon_i \cdot k_j)(k_i  \cdot l)/(k_i \cdot k_j)$ using polarization completeness relations \cite{}. Thus, any $N$-gon tensor integral permits a partial fraction decomposition in terms of inverse propagators, and thus can be mapped to a scalar basis of integrals:
\be\label{TenRed1loop}
T_{\mu_1\mu_2...\mu_r} I_N^{\mu_1\mu_2...\mu_r} = \sum_{M=1}^N C_M I_M
\ee
where $T_{\mu_1\mu_2...\mu_r}$ and $C_M$ are functions exclusively of external kinematics. This relationship holds $D$-dimensionally for any $r$-rank $N$-gon one-loop integral. Thus, any one-loop amplitude can be expressed completely in terms of a basis of one-loop integrals. This special property of one-loop integration is \textit{not} the case for generic loop order, where the basis of ISPs grows faster than number of inverse propagators. As such, any universal integral basis beyond one-loop must include integrals with non-unit exponents for the propagators.

The remarkable simplicity of one-loop amplitudes is even more dramatic in a fixed spacetime dimension. In $D=4$, the kinematic projection to a basis of scalar integrals saturates at the box integral due to the appearance of Gram determinants. Roughly speaking, $N$-gon integral coefficients of \eqn{TenRed1loop} come dressed with a factor of $\mathcal{G}_N$,
\be
\mathcal{G}_N = \text{det} (k_i \cdot k_j)
\ee
where $k_i$ are the external momenta flowing into the $N$-gon integral. In fixed spacetime dimension, $D$, the momenta are $D$-component vectors and thus the Lorentz product matrix $(k_i \cdot k_j)$ must have a null space for $N>D$. Due to this physical constraint, $C_N \neq 0$, only when $N\leq D$. We will demonstrate how this can be used in 4D integral construction in the next section.

@@

The existence of closed for expressions, like those in \eqn{eq:integrals}, while common for one-loop integrals, and incredibly rare for higher loop order outside of very specialized kinematic regimes. Luckily, in this work, the integral basis at four-poinnt two-loop order for even-point theories can be reconstructed from the one-loop expressions in \eqn{eq:integrals}. This is a special property of recursively one-loop amplitudes, which we will describe in detail in \sect{sec:recInt}
\subsection{On-shell Unitarity methods}\label{sec:genU}
Above we have reviewed the known behavior of one-loop integral bases in $D=4$ spacetime dimensions. In this section we will briefly demonstrate why traditional 4D unitarity methods are so powerful -- and why we must go beyond them to accommodate the multi-loop calculations of interest in this work. 
\paragraph{Integrands from 4D Cuts}
Applying one-loop integral reduction in 4D tells us that any one-loop massless amplitude, for which tadpoles are scaleless, can be expressed in terms of box, triangle, and bubble diagrams, with each integral weighted by dimensionally-dependent functions of kinematics:
\be\label{eq:4DUn}
\mathcal{A}^{\text{1-loop}}= \sum_{N=2}^4 C_N(D)I_N^D = \sum_{N=2}^4 C_N(4)I_N^D+ \mathcal{R}
\ee
where the scalar basis-integrals, $I_N^D$, are evaluated in $D=4-2\epsilon$, and $\mathcal{R}$ is a rational term that emerges from series expanding around $\epsilon=(4-D)/2$. To construct integral coefficients, $C_N$, we make use of the Optical Theorem for the $S$-matrix, which states that,
\be
S^\dagger S = \mathbb{I} \quad \Rightarrow \quad 2\,\text{Im}(T) = T^\dagger T
\ee
where $S=\mathbb{I}+iT$. Applying relation to \eqn{eq:4DUn} allows us to identify the imaginary part of $\mathcal{A}^{\text{1-loop}}$ with the imaginary parts of each basis integral weighted by the respective 4D coefficient, $C_N(4)$:
\be\label{eq:UnMeth1loop}
\text{Im}(\mathcal{A}^{\text{1-loop}}) = \sum_{N=2}^4 C_N(4)\, \text{Im}(I_N^D)
\ee
Application of the optical theorem to constrain on-shell amplitudes is known as the unitarity method \cite{}. In order to isolate each $C_N(4)$ appearing in \eqn{eq:4DUn}, we must employ \textit{generalized unitarity}, which places particular internal legs on-shell. That is, for each internal loop propagator, $1/P_i^2$, that we wish to take on-shell, we make the following replacement inside the integral:
\be
\frac{i}{P_i^2 + i\delta} \rightarrow 2\pi \delta(P_i^2)
\ee
This replacement of the internal propagators will select out the requisite imaginary parts needed to determine the integral coefficients. This approach to one-loop amplitude construction was systematized in ref. \cite{}, which demonstrated how to directly extract the integrand coefficients directly from 4D cuts:
\begin{align}
C^{(P_iP_jP_k)}_4 &\sim \boxCut
\\
C^{(P_iP_j)}_3 &\sim \triCut - \sum_{k} C^{(P_iP_jP_k)}_4
\\
C^{(P_i)}_2 &\sim \bubCut- \sum_{j} C^{(P_iP_j)} - \sum_{j,k} C^{(P_iP_jP_k)}_4
\end{align}
where exposed legs are summed over all internal states crossing the cut, and the labels $P_i$ are external momentum scales flowing into the specified cuts. The subtractions account for the overlapping information in the respective $N$-gon cuts. By momentum conservation, the momentum flow in the unlabeled vertex is completely determined by the other $N-1$ vertices.

@@

While this method is incredible powerful, as we can see from \eqn{eq:UnMeth1loop} it misses a key piece of the amplitude -- namely, the rational terms, $\mathcal{R}$. For some theories like $\mathcal{N}=4$ super-Yang-Mills in $D=4$ \cite{}, rational terms are absent because the loop power counting for an $N$-gon cut is bounded to be less than $N-2$. However, for generic theories of interest in this paper, like Born-Infeld, rational terms are present and necessary for extracting anomalous matrix elements and higher-loop sub-divergences. While there are methods developed to extract these rational terms at one-loop using $\mu$-integrals \cite{}, at generic loop order evaluating these $\mu$-integrals becomes increasingly complicated. For this reason, to perform the multi-loop calculations in this paper, we will employ cut construction in general dimensions \cite{Bern:1995db,Bern:1996ja}. 

@@

In this section we will describe the methods we have developed to compute the multi-loop results of this paper. We begin by introducing the notion of \textit{Even-point Multi-loop Unitarity} (EMU), which is an organizational principle at the foundation of our unitarity-based integrand construction. EMU is an extension of the method of maximal cuts \cite{}, which is a hierarchical approach to perturbative calculations \cite{}. The method of maximal cuts is an ansatz-based approach, that feeds successively higher-point contact information to an integrand constrained by graph automorphism symmetry, and color-kinematics duality for permissible theories. In contrast, EMU is a \textit{constructive} algorithm aimed at capturing all the perturbative information needed at general loop order and multiplicity. We describe the algorithm below, and provide a 2-loop 4-point example that we choose to study in this paper.

@@

We note that these coefficients are also sufficient to evaluate the contribution from the double-bubble integral, appearing in the bottom left of \Fig{fig:contrib2Loop}. This is best demonstrated with an explicit example. Consider the following integral, $I^{\text{ex}.}_{\text{2bub}}$, that functionally captures terms that could appear in the cut construction of the double-bubble integral at two-loop:
\begin{equation}
I^{\text{ex}.}_{\text{2bub}} \equiv \int \frac{d^D l_{1}d^D l_{2}}{(2\pi)^{2D}} \frac{(l_1\!\cdot\! l_2)^2(l_1\!\cdot\! v_1)(l_2\!\cdot\! v_2)}{l_1^2(l_1+k_{12})^2 l_2^2(l_2+k_{12})^2}
\end{equation}
where $v_i$ is a stand-in for external kinematics, $\varepsilon_i$ or $k_i$. While the numerator mixes factors of $l_1$ and $l_2$, the denominator can be separated. This allows us to re-express the above integral completely in terms of iterated tensor bubbles:
\begin{equation}
I^{\text{ex}.}_{\text{2bub}} = I_{2}^{\alpha\beta\gamma}(s_{12}) I_2^{\mu\nu\rho}(s_{12})\eta_{\alpha\beta}\eta_{\mu\nu}v_{1 \gamma}v_{2 \rho}\,.
\end{equation}
Then, by applying \eqn{tensorReductionBubble}, and plugging in the expressions for $a_{(m,k)}$, the kinematic numerator of $I^{\text{ex}.}_{\text{2bub}}$ no longer mixes loop momenta. Thus, the integral is separable and can be expressed as a product of scalar bubbles integrated over $l_1$ and $l_2$.

A similar procedure can be applied to the top left diagram of \Fig{fig:contrib2Loop}. However, the integration procedure is a little more delicate. Since $I^{\text{ex}.}_{\text{2bub}}$ can be expressed as an iterated bubble, we only needed to consider \textit{integer powers} of the denominators when constructing the recursion relations. In contrast, the integral in the top left of \Fig{fig:contrib2Loop} is \textit{recursively one-loop}, which will lead to \textit{non-integer}, $\epsilon$-dependent powers of loop propagators. Thus, we are interested in performing a tensor reduction on the following triangle integral

>>>>

When expressed in this form, many gauge theories reveal hidden redundancy that is obscured by the color-dressed amplitude formulation of \eqn{fullAmp}. When expressed in this form, the partial amplitudes for large catalog \cite{BCJreview} of gauge theories, including Yang-Mills and NLSM, are linearly dependent. The first such set of partial amplitude relations identified by Kleiss and Kluijf \cite{Kleiss:1988ne}, known formally as KK relations, relates amplitudes different color orderings and signature:
\be\label{KK}
A(1,\alpha,n,\beta)=(-1)^{|\alpha|}\sum_{\sigma\in \alpha \shuffle \beta^T} A(1,\sigma,n)
\ee
where $\beta^T$ denotes the inverse ordering of $\beta$, and $\shuffle$ is the shuffle product that enumerates ordered permutations of $\alpha$ and $\beta^T$. More recently, Bern, Johansson and one of the authors, further identified a set of kinematic relations that reduced the size of $n$-point partial amplitude basis down to $(n-3)!$ via BCJ relations \cite{BCJ}:
\be\label{BCJ}
\sum_{i=2}^{n-1}k_1\cdot(k_2+\cdots +k_i) A(2,...,i,1,...,n)=0
\ee
Underlying both sets of amplitude relations are hidden graphical organization principles that manifest the amplitude level redundancy of \eqn{KK} and \eqn{BCJ}, which we will describe in the next section. Before moving on, we emphasize that everything described thus far applies in arbitrary spacetime dimensions. As such, the aforementioned nomenclature and organization is well suited for constructing loop integrands needed for dimensional regularization.
\label{subsec:AmpReview}

@@

Formal polarizations will always be expressed as $\varepsilon_a^\mu$ to distinguish between the extra dimensional bit of the integration measure in $D=4-2\epsilon$. As noted, it will be convenient to take $D\rightarrow 4$ in certain circumstances after integration. We will use the conventions of ref. \cite{jjmcTASI2014} throughout the text. Mandelstam invariants with these conventions are expressed as follows, 

@@

We could also consider higher derivative generalizations of the NLSM that are invariant under the shift symmetry of \eqn{eq:shiftSym}. Such EFTs are known formally as \textit{chiral perturbation theory} \cite{Weinberg:1978kz,Gasser:1983yg}, and descriptors of pion scattering below the chiral symmetry breaking scale of QCD. 

@@

By construction, each of the higher derivative operators appearing above are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{Adams:2006sv,Manohar:2008tc,Bellazzini:2016xrt,Guerrieri:2020bto}, and soft theorem bootstraps of refs. \cite{Cheung:2015ota,Cheung:2016drk,Low:2019ynd}. Considering the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $SU(2)$ NLSM have been computed through two-loop order \cite{Bijnens:1995yn,Bijnens:1997vq,Girlanda:1997ed} . In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. In doing so, we show that the IR divergences of the theory exponentiates in the large $N_c$ limit. 

@@

The second even-point EFT we will study is Born-Infeld theory \cite{Born:1934gh}. This theory describes the dynamics of open string endpoints with $U(1)$ charge attached to a $D$-dimensional spacetime with Dirichlet boundary conditions. Given the stringy dynamics orthogonal to the spacetime, the electric field generated by the charged endpoints has a maximum field strength of order the inverse string tension, $1/\alpha'$. The effective Lagrangian for this theory

@@

where $\varphi$ is the spacetime coordinate in the orthogonal space. This theory is uniquely determined by requiring the a polynomial EFT of $X=(\partial \varphi)^2$ is invariant under the field redefinition \cite{Cheung:2016drk},

@@

When the transformation law leave the polynomial $P(X)$ theory invariant, there are an additional $D+1$ conserved charges, and the Poincar\'{e} symmetry of $P(X)$ EFT is promoted from $D$ to $(D+1)$ dimensions. One can verify that applying \eqn{DBIShift} to \eqn{DBI} shifts the Lagrangian by a total derivative \cite{deRham:2010eu}. Similar to the soft theorem constraints on the building blocks of NLSM and $\chi \text{PT}$, the DBI action can be bootstrapped by requiring the amplitudes vanish at $\mathcal{O}(p^2)$ when taking external momentum, $p$, soft \cite{Cheung:2016drk}. 

@@

Now we will describe the supersymmetric extensions of the Born-Infeld action given above in \eqn{biLag}. First we start by defining the theory of shift symmetric fermions first written down by Volkov, Akulov (VA) \cite{Volkov:1973ix}. The easiest way to define this shift symmetric theory is by first constructing manifestly symmetric 1-forms, and constructing a volume form in the veirbein frame. That is, define a 1-form, $ \omega^m = e^m_\mu dx^\mu$, where the veirbein (frame fields), $e^m_\mu$, is expressed in terms of the fermion fields, $\lambda$, as follows:
 \be
 e^m_\mu = \delta^m_\mu + i \bar{\lambda} \Gamma^m\! \stackrel{\leftrightarrow}{\partial_\mu}\!  \lambda = \delta^m_\mu + i (\bar{\lambda} \Gamma^m \partial_\mu  \lambda - \partial_\mu \bar{\lambda} \Gamma^m   \lambda)
 \ee
 This frame field is manifestly invariant under the superspace shift:
\be
\delta \lambda \rightarrow \eta \qquad  \delta \bar{\lambda} \rightarrow \bar{\eta} \qquad  \delta x^\mu \rightarrow x^\mu - i (\bar{\lambda} \,\Gamma^\mu \eta- i \bar{\eta} \,\Gamma^\mu \lambda)
\ee
 The Volkov-Akulov action in the veirbein frame is then simply the fermionic world-volume integral over the $D$-form \cite{Kallosh:1997aw},
 \be
 S^{\text{VA}} = \int \omega^1 \wedge \omega^2 \wedge \cdots \wedge \omega^D\,.
 \ee
 This is not dissimilar from the construction of the pure scale DBI action above. Likewise, we can write the Lagrangian in the spacetime coordinate from akin to \eqn{biLag} and \eqn{DBI}. Transforming back into spacetime coordinates, yields the following Lagrangian density for self interaction VA fermions upto four-point interactions:
\begin{equation}
 \label{vaLag}
  \alpha'^{\,2}\mathcal{L}^{\text{VA}} = \text{det}(e^m_\mu) =  i\bar{\lambda}\!\stackrel{\leftrightarrow}{\slashed{\partial}}\!\lambda+ \frac{1}{2}(\bar{\lambda}\Gamma^\mu \partial^\nu \lambda)(\bar{\lambda}\Gamma_\mu \partial_\nu \lambda)+\mathcal{O}(\partial^4\lambda^6) 
  \,.
\end{equation}
 A convenient choice of field redefinition was proposed by Komargodski and Seiberg (KS) \cite{Komargodski:2009rz,Kuzenko:2010ef}, which defines a new fermion field, $\psi = \psi(\lambda,\bar{\lambda})$, in a way that is perturbatively equivalent to the VA fermions. This construction is equivalent to a theory of nilpotent chiral superfields \cite{Rocek:1978nb,Casalbuoni:1988xh,Ferrara:2014kva}, relevant for the construction of $\alpha$-attractor models of inflation \cite{Kallosh:2013yoa}. Applying the KS nonlinear field redefinition yields the considerably simpler Lagrangian for VA theory:
 \be
   \alpha'^{\,2}\mathcal{L}^{\text{VA}} = i\bar{\psi}\slashed{\partial}\psi +\frac{1}{2} \bar{\psi}^2 \partial^2 \psi^2 + \frac{1}{4}\psi^2\bar{\psi}^2 \partial^2 \psi^2\partial^2 \bar{\psi}^2
 \ee
With the world-volume formulation of VA theory in hand, we can also trivially construct the action for maximal $\mathcal{N}=1$ supersymmetric Born-Infeld theory in $D=10$ \cite{Tseytlin:1999dj,Bergshoeff:2013pia}. We simply replace the flat space background of \eqn{biLag} with the fermionic background of \eqn{vaLag},
\be
 \label{DBIVALag}
S^{\text{DBIVA}}_{\mathcal{N}=1} = \int \omega^1 \wedge\omega^2 \wedge \cdots \wedge \omega^{10} \sqrt{\text{det}(\eta_{mn}+ \alpha F_{mn})}
\ee
It is worth noting that the fermion-vector interaction is introduced via the \textit{non-minimal} coupling between the field strength and the fermionic spacetime metric. This differs from the linearly realized supersymmetry of super-Yang-Mills theory, which minimally couples the vector mode to the chiral fermion. This non-minimal coupling is a requirement of supersymmtry due to the even-point nature abelian vector theories like Born-Infeld.

@@

Furthermore, factors of $(\epsilon_i\cdot l)$ can be mapped to tensor structures with factors of $(\epsilon_i \cdot k_j)(k_i  \cdot l)/(k_i \cdot k_j)$ using polarization completeness relations \cite{Bern:2017tuc}. Thus, any $N$-gon tensor integral permits a partial fraction decomposition in terms of inverse propagators, and thus can be mapped to a scalar basis of integrals:
\be\label{TenRed1loop}
T_{\mu_1\mu_2...\mu_r} I_N^{\mu_1\mu_2...\mu_r} = \sum_{M=1}^N C_M I_M
\ee
where $T_{\mu_1\mu_2...\mu_r}$ and $C_M$ are functions exclusively of external kinematics. This relationship holds $D$-dimensionally for any $r$-rank $N$-gon one-loop integral. Thus, any one-loop amplitude can be expressed completely in terms of a basis of one-loop integrals. This special property of one-loop integration is \textit{not} the case for generic loop order, where the basis of ISPs grows faster than number of inverse propagators. As such, any universal integral basis beyond one-loop must include integrals with non-unit exponents for the propagators.

The remarkable simplicity of one-loop amplitudes is even more dramatic in a fixed spacetime dimension. In $D=4$, the kinematic projection to a basis of scalar integrals saturates at the box integral due to the appearance of Gram determinants. Roughly speaking, $N$-gon integral coefficients of \eqn{TenRed1loop} come dressed with a Gram determinant factor of $\mathcal{G}_N$,
\be
\mathcal{G}_N = \text{det} (k_i \cdot k_j)_{\!N\!\times\! N}
\ee
where $k_i$ are the external momenta flowing into the $N$-gon integral. In fixed spacetime dimension, $D$, the momenta are $D$-component vectors and thus the Lorentz product matrix $(k_i \cdot k_j)_{\!N\!\times\! N}$ must have a null space for $N>D$. Due to this physical constraint, $C_N \neq 0$, only when $N\leq D$ \cite{Bern:1993kr}. We will demonstrate how this can be used in 4D integral construction in the next section.

@@

The existence of closed form expressions, like those in \eqn{eq:integrals}, while common for one-loop integrals, are incredibly rare for higher loop order outside of very specialized kinematic regimes. Luckily, in this work, the integral basis at four-poinnt two-loop order for even-point theories can be reconstructed from the one-loop expressions in \eqn{eq:integrals}. This is a special property of recursively one-loop amplitudes, which we will describe in detail in \sect{sec:recInt}.
\subsection{On-shell Unitarity methods}\label{sec:genU}
Above we have reviewed the known behavior of one-loop integral bases in $D=4$ spacetime dimensions. In this section we will briefly demonstrate why traditional 4D unitarity methods are so powerful -- and why we must go beyond them to accommodate the multi-loop calculations of interest in this work. 
\paragraph{Integrands from 4D Cuts}
Applying one-loop integral reduction in 4D tells us that any one-loop massless amplitude, for which tadpoles are scaleless, can be expressed in terms of box, triangle, and bubble diagrams, with each integral weighted by dimensionally-dependent functions of kinematics:
\be\label{eq:4DUn}
\mathcal{A}^{\text{1-loop}}= \sum_{N=2}^4 C_N(D)I_N^D = \sum_{N=2}^4 C_N(4)I_N^D+ \mathcal{R}
\ee
where the scalar basis-integrals, $I_N^D$, are evaluated in $D=4-2\epsilon$, and $\mathcal{R}$ is a rational term that emerges from series expanding around $\epsilon=(4-D)/2$. To construct integral coefficients, $C_N$, we make use of the Optical Theorem for the $S$-matrix, which states that,
\be
S^\dagger S = \mathbb{I} \quad \Rightarrow \quad 2\,\text{Im}(T) = T^\dagger T
\ee
where $S=\mathbb{I}+iT$. Applying relation to \eqn{eq:4DUn} allows us to identify the imaginary part of $\mathcal{A}^{\text{1-loop}}$ with the imaginary parts of each basis integral weighted by the respective 4D coefficient, $C_N(4)$:
\be\label{eq:UnMeth1loop}
\text{Im}(\mathcal{A}^{\text{1-loop}}) = \sum_{N=2}^4 C_N(4)\, \text{Im}(I_N^D)
\ee
Application of the optical theorem to constrain on-shell amplitudes is known as the unitarity method \cite{UnitarityMethod, BDKUniarityReview}. In order to isolate each $C_N(4)$ appearing in \eqn{eq:4DUn}, we must employ \textit{generalized unitarity}, which places particular internal legs on-shell. That is, for each internal loop propagator, $1/P_i^2$, that we wish to take on-shell, we make the following replacement inside the integral:
\be
\frac{i}{P_i^2 + i\delta} \rightarrow 2\pi \delta(P_i^2)
\ee
This replacement of the internal propagators will select out the requisite imaginary parts needed to determine the integral coefficients. This approach to one-loop amplitude construction was systematized in ref. \cite{Forde:2007mi}, which demonstrated how to directly extract the integrand coefficients directly from 4D cuts:
\begin{align}
C^{(P_iP_jP_k)}_4 &\sim \boxCut
\\
C^{(P_iP_j)}_3 &\sim \triCut - \sum_{k} C^{(P_iP_jP_k)}_4
\\
C^{(P_i)}_2 &\sim \bubCut- \sum_{j} C^{(P_iP_j)} - \sum_{j,k} C^{(P_iP_jP_k)}_4
\end{align}
where exposed legs are summed over all internal states crossing the cut, and the labels $P_i$ are external momentum scales flowing into the specified cuts. The subtractions account for the overlapping information in the respective $N$-gon cuts. By momentum conservation, the momentum flow in the unlabeled vertex is completely determined by the other $N-1$ vertices.

@@

While this method is incredible powerful, as we can see from \eqn{eq:UnMeth1loop} it misses a key piece of the amplitude -- namely, the rational terms, $\mathcal{R}$. For some theories like $\mathcal{N}=4$ super-Yang-Mills in $D=4$ \cite{Fusing}, rational terms are absent because the loop power counting for an $N$-gon cut is bounded to be less than $N-2$. However, for generic theories of interest in this paper, like Born-Infeld, rational terms are present and necessary for extracting anomalous matrix elements and higher-loop sub-divergences. While there are methods developed to extract these rational terms at one-loop using $\mu$-integrals \cite{Badger:2008cm}, at generic loop order evaluating these $\mu$-integrals becomes increasingly complicated. For this reason, to perform the multi-loop calculations in this paper, we will employ cut construction in general dimensions \cite{Bern:1995db,Bern:1996ja}. 

@@

In this section we will describe the methods we have developed to compute the multi-loop results of this paper. We begin by introducing the notion of \textit{Even-point Multi-loop Unitarity} (EMU), which is an organizational principle at the foundation of our unitarity-based integrand construction. EMU is an extension of the method of maximal cuts \cite{Bern:2007ct}, which is a hierarchical approach to perturbative calculations \cite{Carrasco:2021bmu}. The method of maximal cuts is an ansatz-based approach, that feeds successively higher-point contact information to an integrand constrained by graph automorphism symmetry, and color-kinematics duality for permissible theories. In contrast, EMU is a \textit{constructive} algorithm aimed at capturing all the perturbative information needed at general loop order and multiplicity. We describe the algorithm below, and provide a 2-loop 4-point example that we choose to study in this paper.

@@

We note that these coefficients are also sufficient to evaluate the contribution from the double-bubble integral, appearing on the right hand side of \Fig{fig:contrib2Loop}. This is best demonstrated with an explicit example. Consider the following integral, $I^{\text{ex}.}_{\text{2bub}}$, that functionally captures terms that could appear in the cut construction of the double-bubble integral at two-loop:
\begin{equation}
I^{\text{ex}.}_{\text{2bub}} \equiv \int \frac{d^D l_{1}d^D l_{2}}{(2\pi)^{2D}} \frac{(l_1\!\cdot\! l_2)^2(l_1\!\cdot\! v_1)(l_2\!\cdot\! v_2)}{l_1^2(l_1+k_{12})^2 l_2^2(l_2+k_{12})^2}
\end{equation}
where $v_i$ is a stand-in for external kinematics, $\varepsilon_i$ or $k_i$. While the numerator mixes factors of $l_1$ and $l_2$, the denominator can be separated. This allows us to re-express the above integral completely in terms of iterated tensor bubbles:
\begin{equation}
I^{\text{ex}.}_{\text{2bub}} = I_{2}^{\alpha\beta\gamma}(s_{12}) I_2^{\mu\nu\rho}(s_{12})\eta_{\alpha\beta}\eta_{\mu\nu}v_{1 \gamma}v_{2 \rho}\,.
\end{equation}
Then, by applying \eqn{tensorReductionBubble}, and plugging in the expressions for $a_{(m,k)}$, the kinematic numerator of $I^{\text{ex}.}_{\text{2bub}}$ no longer mixes loop momenta. Thus, the integral is separable and can be expressed as a product of scalar bubbles integrated over $l_1$ and $l_2$.

A similar procedure can be applied to the first diagram of \Fig{fig:contrib2Loop}. However, the integration procedure is a little more delicate. Since $I^{\text{ex}.}_{\text{2bub}}$ can be expressed as an iterated bubble, we only needed to consider \textit{integer powers} of the denominators when constructing the recursion relations. In contrast, the first integral in \Fig{fig:contrib2Loop} is \textit{recursively one-loop}, which will lead to \textit{non-integer} $\epsilon$-dependent powers of loop propagators. Thus, we are interested in performing a tensor reduction on the following triangle integral
